---
title: "CSDE 502 Winter 2021, Assignment 10"
author: "Courtney Allen; uw id: ckallen"
output: 
    bookdown::html_document2:
        number_sections: true
        self_contained: true
        code_folding: hide
        toc: true
        toc_float:
            collapsed: true
            smooth_scroll: false
    pdf_document:
        number_sections: true
        toc: true
        fig_cap: yes
        keep_tex: yes
urlcolor: blue 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(ggplot2)
library(dplyr)
library(tidyverse)
library(ggthemes)
library(scales)
library(knitr)

```



```{r}
dat <- read.table("C:/Users/court/Google Drive/Yr 1 2021 Wi/CSDE 502 ProSeminar/Homework/Population.txt", skip = 2, header = TRUE)
```


## Introduction

Taiwan, a country in east Asia, is one of the most densely populated countries in the world. 

## Methods

These data will be recategorized to better understand the age structure of the population. 

Three large age groups will be constructed to categorize

* **children** age 0-19
* **adults**   age 20-69
* **elderly**  age 70+

I will look at how the population structure has changed over time and whether the average number of males and females in each of these age groups has changed over time.

### Data

This dataset includes one record for each single year age group (from age 0 to age 99) of the population by year from `r min(dat$Year)` to `r max(dat$Year)`. Each record has the population size for female, males, and total population.

### Analysis

```{r}

# create 10 year age groups
dat <- dat %>% 
  mutate(agegroup = case_when(Age >= 70 ~ '70+',
                              Age >= 60  & Age <= 69 ~ '60-69',
                              Age >= 50  & Age <= 59 ~ '50-59',
                              Age >= 40  & Age <= 49 ~ '40-49',
                              Age >= 30  & Age <= 39 ~ '30-39',
                              Age >= 20  & Age <= 29 ~ '20-29',
                              Age >= 10  & Age <= 19 ~ '10-19',
                              TRUE ~ '0-9'))

dat <- dat %>% 
  mutate(agecat = case_when(Age >= 70 ~ 'Elderly',
                              Age >= 20  & Age <= 69 ~ 'Adults',
                            TRUE ~ 'Children'))
  
male_data <- dat %>% select(c("Year", "Age", "agegroup", "agecat", "Male"))
female_data <- dat %>% select(c("Year", "Age", "agegroup", "agecat", "Female"))



```

## Results

The figure below shows that the population has grown from less than 15,000,000 to over 20,000,000 since 1970. Furthermore, the share of the population that are children has gradually decreased over the years and an overwhelming majority of the population is in the adult age range.


```{r}
dat %>%
  mutate(agecat = fct_relevel(agecat, 
            "children", "adult", "elderly")) %>%
  ggplot(aes(x = Year, y = Total, fill = agecat)) +
  geom_bar(position = "stack", stat = "identity") +
  xlab("Age") +
  ylab("Population") +
  ggtitle("Population by Age") +
  scale_y_continuous(labels=comma) + 
  scale_x_discrete(breaks = seq(1970,2020,5),guide = guide_axis(n.dodge = 2)) + 
  theme(axis.text.x=element_text(angle=90)) +
  theme_wsj() +
  theme(legend.position = "top", legend.title = element_blank())

  
```


```{r}

dat2 <- female_data %>%
  filter(Year==1970 | Year==2020) %>%
  group_by(Year, agecat) %>%
  summarise('avg. number of females' = mean(Female,na.rm=T)) 



dat3 <- male_data %>%
  filter(Year==1970 | Year==2020) %>%
  group_by(Year, agecat) %>%
  summarise('avg. number of males' = mean(Male,na.rm=T)) 
  

```
The table below shows that the number of adult females has doubled since 1970, although the number of female children has halved since then. There are always more females than males in the elderly group regardless of year.

```{r}
kable(dat2)
kable(dat3)
```


## Discussion

This write-up showed the rapid growth of the population of Taiwan and the changing age structure. The population has become elderly and there the share of the population that is adults has grown immensely. 


