---
title: "CSDE 502 HW 4"
author: "Courtney K Allen"
date: "02/03/2021"
output: 
  bookdown::html_document2

---

## practice edit for github


```{r data_vector, , echo=FALSE}
gradpain <- c(rep(0,11), rep(1,1), rep(2,2), rep(3,6), rep(4,8), rep(5,10),rep(6,8), rep(7,30), rep(8,10), rep(9,6), rep(10,2))
```


```{r sample descriptions, echo=FALSE}
gradpain_N <- length(gradpain)
gradpain_mean <- mean(gradpain)
```

# Question 1

The sample has `r gradpain_N` observations.


# Question 2

The mean of the sample is `r gradpain_mean`


```{r box1, echo=FALSE}
function_1 <- function(x) {
  n <- length(x)
  data_sample <- sample(x, size = n, replace = TRUE)
  sample_mean <- mean(data_sample)
  function_list <- c(n,sample_mean)
  print(function_list)
}
  
d.vec <- gradpain
function_1(d.vec) 
  
```

```{r box2, echo=FALSE}

set.seed(7)
hundred_means <- NULL

# function above doesn't seem to save any objects so just repeat code anyway
for (i in 1:100) {
  function_1(gradpain)
  n = length(gradpain)
  data_sample <- sample(gradpain, size = n, replace = TRUE)
  sample_mean <- mean(data_sample)
  hundred_means <- c(hundred_means,sample_mean)
}
  
```


# Question 3

The mean of these hundred means is `r round(mean(hundred_means),2)`.

# Question 4

The standard deviation of these hundred means is `r round(sd(hundred_means),2)`.


```{r box3, echo=FALSE}

mean_list <- NULL
set.seed(7)
function_2 <- function(d.vec, n, num.samples) {
  for (i in 1:num.samples) {
  data_sample <<- sample(d.vec, size = n, replace = TRUE)
  sample_size <<- length(data_sample)
  sample_num  <<- i
  sample_mean <<- mean(data_sample)
  mean_vector <<- c(mean_list,sample_mean)
  mean_means  <<- mean(mean_vector)
  sd_means    <<- sd(mean_vector)
  mean_ci     <<- Rmisc::CI(x = mean_vector, ci = 0.95)
  
  return(function_list2 <- c(sample_size, sample_num, sample_mean, mean_vector, mean_means, sd_means, mean_ci))
  }
}

d.vec <- gradpain
n <- length(gradpain)
num.samples <- 100

function_2(d.vec, n, num.samples)


```


# Question 5

The mean of all the means is `r round(mean_means,2)`.


# Question 6

The standard deviation of all the means is `r round(sd_means,2)`. 


# Question 7

The 95% CI around the mean is `r round(mean_ci,2)`.



# Appendix: All code 


```{r ref.label=knitr::all_labels(), echo=TRUE, eval=FALSE}
```

